---
title: "hw3.Rmd"
author: "Juan Simons, Github: Giltbuckle3"
date: "2/17/2020"
output: html_document
---

```{r}
data(tips, package="reshape2")
library(ggplot2)
```

#Question 1
There are a total of 244 rows, 4 of which contain a size of 1. With my understanding, a party requires more than a single person, therefore, the waitor served a total of 240 parties.
```{r}
parties <- subset(tips, size== 1, select = c(size))
numParty <- 240
numParty
```

#Question 2
Most of the variables are of the integer type, except *total_bill* and *tip*, these have the type of double. The reason being is that these 2 variables represent mony, which includes dollar and cent values. The rest of the variables represent simpler values.
```{r}
typeof(tips$total_bill)
typeof(tips$tip)
typeof(tips$sex)
typeof(tips$smoker)
typeof(tips$day)
typeof(tips$time)
typeof(tips$size)
```

#Question 3
```{r}
day<- levels(factor(tips$day, levels=c('Thur','Fri','Sat','Sun')))
day
```

#Question 4
```{r}
female5<-tips[tips$sex =='Female' & tips$size >= 5, ]
female5
```

#Question 5
As we can see in the chart bellow, there is a relatively linear correlation between the *toal_bill* and *tip*. We see that the higher the amount of the bill, the higher the tip will be.
```{r}
ggplot(aes(x = total_bill, y = tip,
           color='red'), data=tips) + geom_point()
```

#Question 6
There are two really noticeable anomalies in this graph which are far away from most of the group of tips. These points are the one close to *2.5* on the **x** axis and to *5* on the **y** axis; the other is located near *35* on the **x** axis and *1.5* on the **y** axis. The first anomaly we see someone that is tipping more than the average party with the same amount of money spent on the bill. With a little bit of code we see that the party only spent *$7.25* on the bill, yet they decided to give *$5.15* on the tip, they probably felt like making someone's day. The second anomaly we see that a party barely wanted to tip even though the average, compared to his bill amount, gives decent tips. The bill for this case was *$32.83* and the tip was *$1.17*. It is possible they thought that the waiter gave bad service or the food was bad? These situations are quite unpredictable really.
```{r}
bad_Tips<-tips[tips$tip >= 5, ]
bad_Tips
```
```{r}
bad_Tips1<-tips[tips$tip <= 1.5, ]
bad_Tips1
```

#Question 7
As we see below, the average of tips is 2.84.
```{r}
rate<-tips$tip/tips$total_bill-tips$tip
tiprate<-abs(mean(rate))
tiprate
```

#Question 8
The boxplots are very similar to eachother, there is one outlier though, the medians of the plots. On the *Lunch* plot we see that the median is lower than the *Dinner* plot.
```{r}
tiprate <- tips$tip / (tips$total_bill - tips$tip)
boxChart<- ggplot(tips, aes(x=time, y=tiprate)) + 
  geom_boxplot() + scale_x_discrete(limits=c("Lunch", "Dinner")) + labs(x = "Time", y = "Tip rate", title = "Time of Day Vs Tiprate")
boxChart
```

#Question 9
(Note, the graph needs to be full screen in order to visualize data appropiately)
In the graph below we have 2 subgraphs. The first represents the Female sex and the second, the male sex. There are two colors within the lines of the graph. If the line is blue, they are not a smoker, if red, they are a smoker. We see that females who do not smoke tend to pay more tips, tip amount increases whith the *total_amount*, than females who smoke. In the male graph, it is somewhat spread out. We see that the tip amount is higher from smokers up to a certain point in the graph but then drops if the bill amount exceeds that point (which is about **$25**). Once the bill amount is over that, non-smokers tend to tip more.
```{r}
smokeBar<-ggplot(tips, aes(x=total_bill, y=tip, fill=smoker)) +
         theme_bw() +
         geom_bar(position="stack", stat="identity") +
         scale_fill_manual(values=c("blue", "red")) +
         facet_grid(. ~ sex, space="free_x", scales="free_x",
             labeller=label_both)
smokeBar
```

#Question 10
As we see below, females tend to pay more often on Thursdays and Fridays, whereas men, pay more on Saturdays and Sundays. The increase in men paying in the weekend is probably due to men going on dates and then paying for the date and or simply paying for the party.
```{r}
dayBars<- ggplot(tips, aes(day,sex)) +
  geom_jitter(aes(colour = tips$sex)) + labs(x = "Day of the Week", y = "Gender", title = "Gender vs Weekday",colour = "Gender Color")
dayBars
```
